---
title: "Discovering Annotation Resources with _biocdb_"
author: "Martin Morgan"
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Discovering Annotation Resources with biocdb}
  %\VignetteEngine{knitr::rmarkdown}
---

Edited: 22 April, 2016
<br>Compiled: `r BiocStyle::doc_date()`

# Overview

This package provides functionality to easily discover _Bioconductor_
'annotation' packages and UCSC annotation resources. The latter
requires availability of the [RMySQL][1] package, and consequently a
MySQL client.

```{r setup}
suppressPackageStartupMessages({
   library(biocdb)
})
```

The package depends on both [AnnotationDbi][3] and [dplyr][4]. These
packages have some functions that are named the same but with
different arguments and implementation; it is therefore often
necessary to prefix functions with the name of the package, to
disambiguate reference, e.g., `AnnotationDbi::select()`,
`dplyr::select()`.

# Discovering and using _Bioconductor_ annotation packages.

Use `anno()` to discover _Bioconductor_ [annotation][2] packages. The
functions primary argument, `prefix`, is a character(1) regular
expression that restricts displayed packages. Packages are classified
as 'installed' or 'available. The 'installed' packages are installed
in current _R_ libraries (see `.libPaths()`). The 'available' packages
can be installed via `anno()` (described later) or `biocLite()`.

When `prefix` is missing, common annotation packages are
discovered. These include packages whose names start with 'org',
'TxDb', 'FDb', 'BSgenome', 'SNPlocs', as well as 'Homo.sapiens',
'Mus.musculus', and 'Rattus.norvegicus'.

```{r anno-1}
pkgs <- anno()
names(pkgs)
sapply(pkgs, length)
lapply(pkgs, head)
```

Specifying any prefix returns information on all packages with that
prefix, providing an easy way to discover installed and / or available
packages.

```{r anno-TxDb}
anno("TxDb")
```

If the prefix identifies a single package, then that package is
attached to the current _R_ session, or if necessary installed (via
`biocLite()`) and then attached. The return value is the symbol used
to access data in the package or, if there is no such symbol, `NULL`.

```{r anno-org.Hs.eg.db}
db <- anno("org.Hs.eg")     # uniquely identifies org.Hs.eg.db
db
head(AnnotationDbi::keys(db))
```

Many _Bioconductor_ annotation packages are SQLite data bases. The
argument `dplyr=TRUE` applied to these packages returns a dplyr object
representing the data base.

```{r anno-org.Hs.eg.db-dplyr}
anno("TxDb.*dm6", dplyr=TRUE)
```
# Discovering and using UCSC MySQL resources

The `ucsc()` function provides SQL-level access to UCSC SQL tables. It
requires a MySQL client, and the [RMySQL][1] _R_ package. The UCSC
genome browser web site includes [MySQL database descriptions][3]. A
simple illustration retrieves tables for hg38; the return value is a
dplyr database connection.

```{r ucsc-hg38}
db <- ucsc("hg38")
db
```

# About this document

```{r sessionInfo}
sessionInfo()
```

[1]: https://cran.r-projec.org/package=RMySQL
[2]: http://bioconductor.org/packages/release/BiocViews.html#___AnnotationData
[3]: http://genome.ucsc.edu/goldenPath/help/mysql.html
